\centerline{\strong{Copyright \copyright\ 2000 Gregory Trubetskoy All rights reserved.}}

Redistribution and use in source and binary forms, with or without
modification, are permitted provided that the following conditions
are met:

\begin{enumerate}

\item
Redistributions of source code must retain the above copyright
notice, this list of conditions and the following disclaimer. 

\item
Redistributions in binary form must reproduce the above copyright
notice, this list of conditions and the following disclaimer in the
documentation and/or other materials provided with the distribution.

\item
All advertising materials mentioning features or use of this software
must display the following acknowledgment: "This product includes
software developed by Gregory Trubetskoy for use in the mod_python
module for Apache HTTP server (http://www.modpython.org/)."

\item
The names "mod_python", "modpython" or "Gregory Trubetskoy" must not 
be used to endorse or promote products derived from this software 
without prior written permission. For written permission, please 
contact grisha@ispol.com.

\item
Products derived from this software may not be called "mod_python"
or "modpython", nor may "mod_python" or "modpython" appear in their 
names without prior written permission of Gregory Trubetskoy. For 
written permission, please contact grisha@ispol.com..

\item
Redistributions of any form whatsoever must retain the following
acknowledgment:
"This product includes software developed by Gregory Trubetskoy
for use in the mod_python module for Apache HTTP server 
(http://www.modpython.org/)."

\end{enumerate}

THIS SOFTWARE IS PROVIDED BY GREGORY TRUBETSKOY ``AS IS'' AND ANY
EXPRESSED OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR
PURPOSE ARE DISCLAIMED.  IN NO EVENT SHALL GREGORY TRUBETSKOY OR
HIS CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL,
SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT
NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES;
LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION)
HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT,
STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED
OF THE POSSIBILITY OF SUCH DAMAGE.

This software is based on the original concept
as published in the book "Internet Programming with Python"
by Aaron Watters, Guido van Rossum and James C. Ahlstrom, 
1996 M\&T Books, ISBN: 1-55851-484-8. The book and original
software is Copyright 1996 by M\&T Books.

This software consists of an extension to the Apache http server.
More information about Apache may be found at 

http://www.apache.org/

More information on Python language can be found at

http://www.python.org/

